---
title: "Covid91 Vaccine Study Final 2023"
author: "Ashley Davis"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of sub populations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


$H_0$: The COVID vaccine is not effective in treating COVID among subgroups.

$H_A$: The COVID vaccine is effective is treating COVID among subgroups.

# Methods


The statistical analysis well be using is cat~ cat because the covid vaccine should be effective among subgroups.

The method we will use to display the data is a bar chart,cross table, and chi squared test

# Results

## Males

### Descriptive Results
```{r}
maleData = filter (FinalData,sex == "M")
```

#### Graphical Results

```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")

```

Based on the graph, it is showing us from the all male group of people who contracted COVID and the ones who received the vaccine had less symptoms than those that received the placebo, while the group that didn't contract COVID showed no difference between those who were vaccinated and un-vaccinated.

#### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment , data=maleData)
rowPerc(table1)
colPerc(table1)
```

The table shows that the group with covid experience a 2x difference between the vaccinated and the un-vaccinated.The placebo group had a 70% infected rate compared to the vaccinated group which only had 30%.The Ok group had no significant difference between the un-vaccinated and vaccinated


#### Inferental Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p value is 4.631e-14  making it to small to be in significant therefore disproving the null hypothesis

Based on Fishers exact value You become .35% more susceptible to contracting COVID if you are not vaccinated

## Females

```{r}
femaleData = filter (FinalData,sex == "F")
```

### Descriptive Results

#### Graphical Results

```{r}
barchartGC(~infected+treatment,data=femaleData)
barchartGC(~infected+treatment,data=femaleData, type="percent")
```


This graph is showing that from the all female group who contracted COVID, the ones who received the vaccine had less symptoms than those with the placebo while the all the female group that did not contract COVID showed no difference between those who were vaccinated and not vaccinated. 

#### Numerical Results

```{r}
table2 <- xtabs(~infected + treatment , data=femaleData)
rowPerc(table2)
colPerc(table2)
```

The data from this table shows that from the female group that were infected with COVID 19, those who were unvaccinated, 60% of them experienced symptoms while only 40% of the female group who were vaccinated experienced symptoms. It is also showing that within the female placebo group, 2.04% were infected with COVID, while only 1.37% of the vaccinated group contracted COVID.

#### Inferental Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

Based on this chisq test, the p-value is 0.0002732, showing that there is a significant difference, concluding that we are rejecting the null hypothesis. This means that the COVID vaccine is effective within the female subgroup. 

For the fisher exact test show that the odds are about 1.50%, suggesting that there is a significant correlation between the variables, showing that there is strong evidence directing that we are rejecting the null hypothesis. 

## LGBTQ

```{r}
LGBTQData = filter (FinalData,LGBTQ == "gay")
```

### Descriptive Results

#### Graphical Results

```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")
```

This graph is showing that within the LGBTQ subgroup, those who got COVID and got vaccinated experienced significantly more symptoms than those their unvaccinated counterpart. Those who did not get COVID and were both vaccinated and unvaccinated did not show any difference.

#### Numerical Results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```

This table is showing that within the LGBTQ placebo group they experienced 15.38% of symptoms while the the vaccinated group experienced 84.62% of syptoms. It is also showing that 0.66% of the LGBTQ group who had the placebo contracted COVID, while 3.54% of the vaccinated group contracted COVID.

#### Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The chisq test shows that the p-value for the LGBTQ group is 0.0000006668 showing that there is an effect between the COVID vaccine and the LGBT subgroup, concluding that we are rejecting the null hypothesis.  

In the fisher exact test, the percentage odds ratio is about 0.18%, showing that the odds are about 0.18 times lower in one variable group compared to the other, showing a difference between the two groups, which is another example of rejecting the null hypothesis in this group.

## Druggies



### Descriptive Results



#### Graphical Results



#### Numerical Results



#### Inferential Results

# Overall Results and Conclusions